Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-05-29 talk/dls2014/paper/paper.tex
tweaks

We performed all benchmarks on a machine with a Intel Core i7-4770
CPU~@3.40GHz (4 cores, 8 threads).  There are 16~GiB of memory
available and we ran them under Ubuntu 14.04 with a Linux 3.13.0
kernel. The STM system was compiled with a number of segments $N=4$
and a maximum amount of memory of 1.5~GiB (both are configurable at
compile time).

@@

For PyPy-STM the average memory requirement is 21~MiB and there are
$\sim 10$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time. Compared to normal PyPy, we are missing a
memory optimisation to store small objects in a more compact
way. Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
We expect to improve this aspect in the future, in this paper we
want to focus first on performance.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

@@

To asses how well the STM system scales on its own (without any real
workload), we execute the following loop on 1 to 4 threads:
\begin{lstlisting}
def workload():
    i = 20000000
    while i:
        i -= 1
\end{lstlisting}

For the results in figure \ref{fig:scaling}, we averaged
over 5 runs and normalised the average runtimes to the
time it took on a single thread. From this we see that there
is additional overhead introduced by each thread ($13\%$
for all 4 threads together).

>>>>

We performed all benchmarks on a machine with an Intel Core i7-4770
CPU~@3.40GHz (4 cores, 8 threads).  There are 16~GiB of memory
available and we ran them under Ubuntu 14.04 with a Linux 3.13.0
kernel. The STM system was compiled with a number of segments $N=4$
and a maximum amount of memory of 1.5~GiB (both are configurable at
compile time).

@@

For PyPy-STM the average memory requirement is 21~MiB and there are
$\sim 10$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time. Compared to normal PyPy, we are missing a
memory optimisation to store small objects in a more compact
way. Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
We expect to improve this aspect in the future; in this paper we
want to focus first on performance.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

@@

To asses how well the STM system scales on its own (without any real
workload), we execute the following loop on 1 to 4 threads on our
Python interpreter with STM:
\begin{lstlisting}
def workload():
    i = 20000000
    while i:
        i -= 1
\end{lstlisting}

For the results in figure \ref{fig:scaling}, we averaged over 5 runs
and normalised the average runtimes to the time it took on a single
thread. From this we see that there is additional overhead introduced
by each thread (\remi{$13\%$} for all 4 threads together). While this
is not ideal, we think that \remi{$13\%$} are acceptable on four
threads. In terms of throughput, 4 threads have \remi{$3.53\times$}
more iterations per second than a single thread.
